---
title: "Credit Risk Analysis of Mobile Money"
author: "Benjamin Ekow Attabra"
date: "September 2024"
output:
  html_document: default
  pdf_document: default
---
```
## Introduction

This analysis aims to assess credit risk based on mobile money transaction data. Specifically, we will evaluate the cash flow of the user by analyzing inflows (deposits) and outflows (expenditures) and provide insights into potential credit risk factors.

## 1. Data Preparation

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(lubridate)
library(ggplot2)
library(knitr)

# Load the data
transactions <- read_csv("transactions.csv")

# Convert Transaction Date to Date type
transactions$`Transaction Date` <- dmy(transactions$`Transaction Date`)

# View the structure of the data
str(transactions)

# Display the first few rows of the data
head(transactions)
```

## 2. Exploratory Data Analysis (EDA)

### Summary Statistics

```{r summary-stats}
# Calculate summary statistics for inflows and outflows
summary(transactions$Inflows)
summary(transactions$Outflows)

# Summary statistics for Fees and E-Levy
summary(transactions$Fees)
summary(transactions$E.Levy)
```

### Visualizing Transaction Trends

```{r transaction-trends, echo=FALSE}
# Summarize daily inflows and outflows
daily_summary <- transactions %>%
  group_by(`Transaction Date`, Type) %>%
  summarise(Total_Amount = sum(Inflows + Outflows))

# Plotting inflows and outflows over time
ggplot(data = daily_summary, aes(x = `Transaction Date`, y = Total_Amount, color = Type)) +
  geom_line() +
  labs(title = "Daily Inflows and Outflows Over Time",
       x = "Date",
       y = "Amount",
       color = "Transaction Type") +
  theme_minimal()
```

## 3. Credit Risk Metrics

### Cash Flow Analysis

```{r cash-flow-analysis}
# Calculate net cash flow (Inflows - Outflows)
transactions <- transactions %>%
  mutate(Net_Cash_Flow = Inflows - Outflows)

# Summary of net cash flow
summary(transactions$Net_Cash_Flow)

# Plotting net cash flow over time
ggplot(transactions, aes(x = `Transaction Date`, y = Net_Cash_Flow)) +
  geom_line(color = "blue") +
  labs(title = "Net Cash Flow Over Time",
       x = "Date",
       y = "Net Cash Flow") +
  theme_minimal()
```

### Delinquency Risk

```{r delinquency-risk}
# Check for days with negative net cash flow (potential credit risk)
delinquent_days <- transactions %>%
  filter(Net_Cash_Flow < 0)

# Display delinquent days
kable(delinquent_days, caption = "Days with Negative Net Cash Flow")
```

```{r}
# Load the dataset
transactions <- read_csv("transactions.csv")

# Calculate Net Cash Flow (Inflows - Outflows)
transactions <- transactions %>%
  mutate(Net_Cash_Flow = Inflows - Outflows)

# Calculate total net cash flow for the entire period
total_net_cashflow <- sum(transactions$Net_Cash_Flow)

# Determine lending decision based on net cash flow
if (total_net_cashflow > 0) {
  lending_decision <- "Lend"
} else {
  lending_decision <- "Do not lend"
}

# Output the results
cat("The total net cash flow calculated from the mobile money transactions is: ", total_net_cashflow, "\n")
cat("Given the total net cash flow of ", total_net_cashflow, ", the conclusion drawn from this analysis is: '", lending_decision, "'.\n\n")
```

## 4. Conclusion

This report provided an overview of the user's cash flow through mobile money transactions. By examining inflows, outflows, fees, and net cash flow over time, we identified key periods of potential credit risk based on negative cash flow. Further recommendations for improving cash management or extending credit can be made based on these results.
```